\chapter{Player Commands}
\label{app:commands}
{\scriptsize
\begin{longtable}{p{4cm}p{9cm}}
apply $<$string$>$ & Without any parameter, applies the top object in the ground view. With a string, applies the first object matching the string. \\
bind & Used to bind commands to keys. \\
brace & Toggles brace status. While braced you will not move. \\
cast $<$spell$>$ & Readies named spell. \\
clearinfo & Clears the text screen. \\
disarm & Executes a disarming action in facing direction. \\
dm $<$password$>$ & Player becomes the dm. You character is unsavable after this. \\
drop $<$object$>$ & Drops object from player inventory. All items matching $<$object$>$ are dropped, unless they are locked or cursed. \\
dropall &  Drops all but locked objects in inventory. \\
east & Execute a move east. \\
examine $<$object$>$ & Examines object, defaults to top object in look window. \\
get & Picks up top object in look window. \\
gsay & Sends a message only to members of your party.\\
help $<$subject$>$ & Help on subject, defaults to shows list of available help. If $<$subject$>$ is {\tt commands}, will display a list of available commands. \\
hiscore & Show list of highest player scores. \\
inventory & Lists all objects in your inventory in text window. \\
invoke $<$spell$>$ $<$string$>$ & Readies, then casts a spell. String is optional \\
listen & Changes the level of your hearing. \\
mapinfo & Show information about current map. \\
maps $<$name$>$ & Show status of all active maps. If $<$name$>$ is provided, only lists maps with matching name. \\
motd & Show message of the day again. \\
north & Execute a move north. \\
northeast & Ditto to north east. \\
northwest & Ditto to the north west. \\
output-count & Toggles message grouping, so that identical messages are put together. \\
output-sync &  No idea what this does. \\
party $<$command$>$ & Set of commands used with simple party system. \\
peaceful & Toggle peaceful status. \\
pickup $<$command$>$ & Change pickup status by number or value density. \\
prepare & Alias for cast command. \\
quit & Quit the game and totally removes your character. \\
ready\_skill $<$skill$>$ & Prepare a skill for use. \\
rotateshoottype & Rotate the range slot by 1. \\
save & Save your current position and status. \\
savewinpos & Save your current window layout. \\
say & Say something to all players on your map. \\
search & Execute a search in all nearby squares.\\
search-items $<$command$>$ & Change the status of items searched for. \\
shout & Send a message to all players regardless of map.\\
show & \\
showinvicon & Toggle status of invicon in inventory window. \\
skills & Show all available skills, experience categories and level. If a string is provided, only lists matching skills.\\
south & Execute a move south.\\
southeast & Ditto to southeast. \\
southwest & Ditto to southwest. \\
stay & \\
take & Alias for pickup. \\
tell $<$who$>$ $<$msg$>$ & Tell character who the msg. \\
throw & Throws an item. \\
time & Displays the ingame time \\
title $<$string$>$ & Change your title to string.\\
unbind $<$command$>$ & A set of commands to reverse key bindings. \\
use\_skill $<$skill$>$ $<$string$>$ & Ready, then use a skill. \\
version & Print out all the contributors to \cf .\\
west & Execute a move west. \\
who & Show what players are currently logged on. If a parameter is specified, only displays players in regions with matching name. \\
wimpy $<$percent$>$ & Auto-matically run away when hp$<$\%$\times$Max hp. \\
\end{longtable}

Commands for Dungeon Masters only:
\begin{longtable}{p{4cm}p{9cm}}
malloc & Shows which stuff is taking up how much memory. \\
sstable & Sends to the server log statistics about shared strings. \\
strings & Shows the status of shared sting parameters. \\
sync $<$integer$>$ & Change sync value to integer. \\
\end{longtable}
}
