
\chapter{About Characters\index{characters}}

\section{Character Attributes\index{characters, attributes}}\label{sec:char_attr}

Each player interacts in the \cf\ world through the persona of a
character. In turn, the interaction between the character and
the game world is mediated by the ``attributes'' of the character.
After all, the player is not going to play him/herself! In \cf\ we
chose to parameterize the acceptable limits of player behavior
via the values of the character attributes which, in turn, help
to determine the success of any action taken by the player. There
are no limits per se on what you can attempt to
do with your character; rather, the attributes of a character
indicate the certain ``natural'' talents and inclinations.
Two concepts which are relevant to \cf\ character attributes are
``stats'' and ``class''.

\subsection{Character Stats\index{stats}\index{characters, stats}}

Character statistics (or ``stats'' for short) can be
divided into two types: primary and secondary. What's the difference
between them? Secondary stats are calculated from a number of things
including the primary stats. But the reverse isn't true, secondary
stats never have any influence on the primary stats. In playing the game, the
player may find that either of these kinds of stats may be changed
for better or worse. In general, the primary stats change much less
often than the secondary stats. Equipment, magic, and death are
just three examples of the many things which can alter the values of
the character stats. The current values of both the primary and
secondary stats may be viewed in the stat window. Four important
secondary stats\emdash food, grace, hitpoints and mana also appear
again in the stat-bar window.


\subsubsection{Primary stats\index{stats, primary}}
\index{STR}\index{CON}\index{DEX}\index{INT}\index{WIS}\index{CHA}
\index{POW}
\index{strength}\index{dexterity}\index{constitution}
\index{power}\index{intelligence}\index{wisdom}
\index{charisma}

The seven primary stats are:
{\small
\begin{quote}
$\bullet$ {\bf Strength} (``STR'') $-$
a measure of the physical strength. \\
$\bullet$ {\bf Dexterity} (``DEX'') $-$
measures physical agility and speed. \\
$\bullet$ {\bf Constitution} (``CON'') $-$
measures physical health and toughness. \\
$\bullet$ {\bf Intelligence} (``INT'') $-$
measures ability to learn skills and \incantation s. \\
$\bullet$ {\bf Wisdom} (``WIS'') $-$
measures the ability to learn/use \divinemagic . \\
$\bullet$ {\bf Power} (``POW'') $-$
measures magical/spiritual potential. \\
$\bullet$ {\bf Charisma} (``CHA'') $-$
measures social and leadership abilities.
\end{quote}
}
Primary stats have a ``natural'' range
between 0 and $\sim$20. The actual upper limit on each primary stat is
set by the chosen character class (see section \ref{sec:char_cls}). You
can raise your primary stats by drinking potions\index{potions} up to
your class natural limit.

There are plenty of items which give you bonuses to your
stats even {\em beyond} your class limit $-$ swords\index{weapons},
armour\index{armour} and rings\index{rings} to
name the most important. You can also read scrolls\index{scrolls} or cast
spells\index{spells} (\incantation s
or prayers) to temporarily raise your stats.  The ultimate maximum value is 30,
and the player class doesn't matter here.

Some quantitative effects of the primary stats are summarized in table
\ref{tab:pri_eff}. The top row in the column header specifies a particular
bonus (i.e. {\tt HpB}, {\tt MgB}, {\tt AcB}, {\tt DmB}, etc.) while the second row
in the header indicates the stat which is used to calculate the value of
that column. Where no stat appears, one of {\em several} stats may be
used to calculate that value. See later parts of the text (particularly
section \ref{sec:stat_calc}) for further details.

\begin{table}
\begin{center}
\scriptsize
\caption{Selected primary stat bonuses/penalties. \label{tab:pri_eff}}
\index{stats, limits}\index{stats, primary}
\vskip 12pt
\begin{tabular}{|r|r|r|r|r|r|r|r|r|r|}\hline
\snglc{Stat}&\snglc{HpB}&\snglc{MgB\tablenotemark{1}}&\snglc{AcB}&\snglc{DmB}&\snglc{Thaco}&\snglc{Max Carry}&\snglc{Speed}&\snglc{\%learn\tablenotemark{2}}&\sngLc{Buy/Sell} \\
\snglc{ }&\snglc{(CON)}&\snglc{}&\snglc{(DEX)}&\snglc{(STR)}&\snglc{(STR)}&\snglc{(STR)}&\snglc{(DEX)}&\snglc{}&\sngLc{(CHA)} \\ \hline\hline
\input{bonus}
\hline
\end{tabular}
\end{center}
\tablenotetext{1}{Either POW or WIS can be used to calculate MgB (magic bonus).}
\tablenotetext{2}{Percentage for learning either skills (INT), \incantation s (INT) or prayers (WIS).}
\end{table}


\subsubsection{Secondary stats\index{stats, secondary}}

The secondary stats are : \\

$\bullet$ {\bf\tt score}\index{score} \emdash\ The total accumulated
experience\index{experience} of the character. {\tt score} is increased
as a reward for appropriate player action and may decrease as a
result of a magical attack or character death (see section
\ref{sec:death} for more about death). The {\tt score} starts at a value
of 0. \\

$\bullet$ {\bf\tt level}\index{level, overall}\index{level} \emdash\
A rating of overall ability whose value is determined from the {\tt score}.
As the {\tt level} of the character increases,
the character becomes able to succeed at more difficult tasks. {\tt level}
starts at a value of 0 and may range up beyond 100. The value of the
stat which appears in the stat window is sometimes known as the {\em overall}
{\tt level}. See section \ref{sec:experience} for more details. \\

$\bullet$ {\bf hit points} (``{\tt Hp}'')\index{stats, Hp}
\index{hit points} \emdash\ Measures of how much
damage the player can take before dying. Hit points are determined from
the player {\tt level} and are influenced by the value of the character CON
(see section \ref{sec:stat_calc}).  {\tt Hp} value may range between 1
to beyond 500 and higher values indicate a greater ability to
withstand punishment. \\

$\bullet$ {\bf mana} (``{\tt Sp}'')\index{stats, mana}
\index{mana}\index{Sp} \emdash\ Measures of how much ``fuel'' the player
has for casting \incantation s.  Mana is calculated from the character
{\tt level}
and the value of the character POW (see section \ref{sec:stat_calc}).
Mana values can range between 1 to beyond
500. Higher values indicate greater amounts of mana. \\

$\bullet$ {\bf grace} (``{\tt Gr}'')\index{Gr}\index{stats, grace}
\index{grace} \emdash\
How favored the character is by the gods. In game terms, how much
\divinemagic\ a character can cast. Character {\tt level}, WIS and POW effect
what the value of grace is (see section
\ref{sec:stat_calc}).  During play, grace values
{\em may} exceed the character maximum; grace can take on large positive
and negative values. Positive values indicate favor by the gods.\\

$\bullet$ {\bf weapon class}\index{weapon class} (``{\tt Wc}'')
\index{stats, Wc}\index{Wc} \emdash\
How skilled the characters melee/missile attack is. Lower values indicate a
more potent, skilled attack. Current weapon, user {\tt level} and STR are some
things which effect the value of {\tt Wc}. The value of {\tt Wc} may
range between 25 and -72.
See section \ref{sec:stat_calc} for a more detailed explanation of weapon
class. See section \ref{sec:combat} to see how {\tt Wc} works in attacking.\\

$\bullet$ {\bf damage}\index{damage} (``{\tt Dam}'')\index{stats, Dam}
\index{Dam} \emdash\ How much
damage a melee/missile attack by the character will inflict. Higher values
indicate a greater amount of damage will be inflicted with each attack.
See section \ref{sec:stat_calc} for a calculation of the character {\tt Dam}.
\\

$\bullet$ {\bf armour class}\index{stats, Ac}\index{armour class}
(``{\tt Ac}'')\index{Ac} \emdash\ How
protected the character is from being hit by any attack. Lower values
are better. {\tt Ac} is based on the character class (table \ref{tab:char_cls})
and is modified by the DEX ({\tt AcB} column in table \ref{tab:pri_eff}) and
current armour worn. For characters that cannot wear armour, {\tt Ac} improves
as their level increases (see section \ref{sec:stat_calc}). \\

$\bullet$ {\bf armour}\index{armour} (``{\tt Arm}'')\index{stats, Arm}
\emdash\
How much damage will be subtracted from successful hits made upon
the character. This value ranges between 0 to 99\%. Current armour worn
primarily determines {\tt Arm} value. \\

$\bullet$ {\bf\tt speed}\index{stats, speed}\index{speed} \emdash\
How fast the player may move.
The value of {\tt speed} may range between nearly 0 (``very slow'') to
higher than 5
(``lightning fast''). Base {\tt speed} is determined from the DEX and modified
downward proportionally by the amount of weight carried which {\em exceeds} the
{\tt Max Carry} limit (table \ref{tab:pri_eff}). The armour worn also sets the
upper limit on {\tt speed} (see the \spoiler\ for these limits). \\

$\bullet$ {\bf weapon speed}\index{stats, weapon speed}\index{weapon speed}
\emdash\ Appears in
parentheses after the {\tt speed} in the stat window. This quantity is
how many attacks the player may make per unit of time.
Higher values indicate faster attack speed. Current weapon and user
DEX effect the value of weapon speed. See section \ref{sec:stat_calc}
for a calculation of weapon speed. \\

$\bullet$ {\bf\tt food} $-$ How full the character's stomach is.
Ranges between 0 (starving) and 999 (satiated).  At a value of 0 the
character begins to die. Some magic can speed up or slow down the
character digestion. Healing wounds will speed up digestion too. \\


\subsection{Character Classes}\index{characters, classes}\label{sec:char_cls}

Much like the older ``paper and pencil'' role-playing games
\cf\ has adopted the idea of character ``class''.
Each class is meant to be a template of a particular ``style'' of play;
therefore each choice of class modifies both the starting
values and sets the natural upper limit on the primary stats.
{\em Important note:} character class is chosen at the time a
character is created and can't be changed later on.
\begin{table}
\begin{center}
\scriptsize
\caption{\cf\ character classes. \label{tab:char_cls}}
\index{characters, classes}
\vskip 12pt
\begin{tabular}{|c|c|l|l|l|l|l|l|l|p{4cm}|}
\hline
Type& &         Str&    Dex&    Con&    Int&    Wis&    Pow& Cha & Special\\
\hline
\hline
\input char.tex
\hline
\end{tabular}
\end{center}
\end{table}

Table \ref{tab:char_cls} shows the various available character classes with
the natural stat limits for each.
Under the ``special'' column several bits of information are included. ``{\tt Ac}''
\index{Ac}\index{stats, Ac} indicates the base armour class
for the character; ``damage''\index{stats, Dam}
indicates the base {\tt Dam}\index{Dam} value.
Some character classes have special attack abilities and certain vulnerabilities,
protections from, and immunities to various attacktypes. Read section
\ref{sec:combat} for more information about what effect these can have.

\subsubsection{Sizing up the character classes}

Generally, the titles of the character classes speak for themselves. But
you can get a better idea of the potential of a class by checking out a
few things. Take a look at the starting equipment
(table \ref{tab:start_equip}) and, if you are using \cf\ compiled with
the skills system, checkout the starting skills for those classes your interested
in (table \ref{tab:skill_start}).


\begin{table}
\begin{center}
\scriptsize
\caption{Starting equipment by character class. \label{tab:start_equip}}
\index{equipment, starting}
\vskip 12pt
\begin{tabular}{|l|l|} \hline
Type & Starting Equipment \\ \hline\hline
\input equip.tex
\hline
\end{tabular}
\end{center}
\end{table}

\begin{table}
\footnotesize
\caption{Starting skills} \label{tab:skill_start}\index{skills, starting}
\vskip 12pt
\begin{center}
\begin{tabular}{|l|l|} \hline
Character class 	& Additional skills\tablenotemark{1} \\ \hline\hline
\input{skills}
\hline
\end{tabular}
\tablenotetext{1}{All character classes start with the skills {\tt melee weapons},
{\tt find traps}, {\tt use magic item}, {\tt literacy}, and {\tt disarm traps}.}
\end{center}
\end{table}

Finally, here are some notes concerning a couple of the more ``exotic'' classes:\\

\noindent{$\bullet$} {\bf Fireborn}\index{Fireborn} \\
{\sl Attacks:} fire, physical \\
{\sl Protections:} immune: fire, poison; vulnerable: ghosthit, drain, cold \\
{\sl Special:} fly, no armour, no weapons. {\tt Ac} 0 \\
Fireborns are fire spirits. They're closely in tune with
magic and learn all types of magic easily.
Being fire spirits, they are
immune to fire and poison, and vulnerable to cold. They are vulnerable to
ghosthit and drain because being mostly non-physical, anything which strikes
directly at the spirit hits them harder. \\

\noindent{$\bullet$} {\bf Monk}\index{Monk} \\
{\sl Attacks:} physical \\
{\sl Protections:} none \\
{\sl Special:} no weapons. \\
Monks are members of various martial arts orders. They have devoted themselves to
a life of contemplation and revelation though physical fighting! Their
life-long secret vows keep them from using all hand-held melee weapons,
but in return they are allowed to learn secret techniques of meditation. \\

\noindent{$\bullet$} {\bf Quetzalcoatl}\index{Quetzalcoatl} \\
{\sl Attacks:} physical \\
{\sl Protections:} immune: fire; vulnerable: paralyze, poison, cold \\
{\sl Special:} no armour, {\tt Ac} 5 \\
Quetzalcoatls are an odd mixture of magic and combat abilities. They are
born knowing the spell of burning hands (heh, its their ``dragon breath'').
But because of their low natural intelligence/wisdom, they have a very hard
time learning new spells. All the same, they may become potent
wizards/priests as they have the highest power bonus, and they will
typically have a large amount of mana and a fair amount of grace.
Quetzalcoatls can be very
devastating fighters at low level. A combination of  their low natural {\tt Ac}
and high base damage tend to make mincemeat
out of low-level monsters. However, at mid-level, they really begin to have
problems because they cannot use armour. \\

\noindent{$\bullet$} {\bf Wraith}\index{Wraith} \\
{\sl Attacks:} cold, physical \\
{\sl Protections:} immune: drain, ghosthit; protected: physical, cold; vulnerable: fire \\
{\sl Special:} {\tt Ac} 6 \\
The Wraith is a creature of the undead. These characters represent revenging spirits come
back to life to work their unholy will on the living. Their undead nature
makes them immune to life-damaging forces and their etherealness protects
them from physical harm. Like all undead, they succumb to fire readily.

\section{Generating a character}\index{characters, generation}\label{sec:char_gen}

When you start up \cf , you will be asked for a character name
followed by a password. If you are playing for the first time, type
any name you like; this will be your character name for the rest of
the life of that character. Next, type in any password.
{\em Remember!:}
you will need to type the same password again to play that character
again! One more note: its not a good idea to use ``real'' passwords
to your computer accounts! Doing so may make your system vulnerable
to unscrupulous \cf\ server administrators.

Your next step will be to generate random (primary) stats for your
character.
You aren't limited to the number of times you can roll your stats\emdash
so have fun.
Notice that the stats are always arranged that the STR stat has
the highest value,
the CHA stat has the lowest. You may re-arrange the order of these values
when you decide that you have rolled some decent stats.
{\em Two points:} you can
never roll a character with better stats than an average of straight 15's,
and you can't roll higher than 18 in a stat.

When you roll your character, the stats displayed are the stats you
will get as a human (which are unmodified).
When satisfied, you can step through a number
of classes, each with special bonuses in stats.

Table \ref{tab:char_cls} shows how your basic stats will
be changed by choosing a different class.
The difference between the natural stat limit and 20 indicates the
bonus/penalty assigned to rolled primary stats. For example, a
barbarian has a maximum strength which is 4 higher than 20$-$that
means he will begin with an additional 4 points added to his
strength roll. On the other hand, a barbarian can never get above 14
in intelligence.  This means that your rolled character will have 6
less in intelligence if you choose that class.  It also means that you
can't be a barbarian if you roll less than 7 in intelligence; the
poor barbarian would have had a stat below 1.

\subsubsection{Selecting a character}

While each class has its particular strengths and weaknesses, in summary
its just a fact that some classes are easier to play than others.

For beginning players, the ``simple fighter'' characters are the easiest to
play successfully. The Dwarf, Human, and Warrior are among good earlier
choices. As you gain experience
with using \cf\ you may wish to branch out into other ``fighter'' characters
such as the Thief or Elf, or try your hand at playing ``spellcaster''
characters like the Wizard, Mage, Cleric or Priest.
The ``exotic'' classes (e.g. Fireborn, Monk, Quetzalcoatl, and Wraith) are
the hardest classes to master.

One more note: If you are using the default game (compiled with the skills
system),
you will probably find any character class that has a low natural INT (for
example, the Barbarian class) will have a fairly difficult time at higher
levels. Skills, and most importantly, spells, will be more difficult
learn. Plan to spend a lot of loot on spell books (\wizbook s and prayer books)
and skill scrolls.
